### 2020-04-20 experiments.tex
ok

Our primary experimental results are a set of comparisons of tools using our method, for three languages: Solidity (the most popular language for smart contracts), Java, and Python.  We used the Universal Mutator tool for all experiments, since e.g., PIT \cite{pitmut} does not produce source-level mutants and the PMD tool requires source code; we also wanted to use a consistent mutation generation method for all languages, to the extent possible; for Solidity and Python, we believe the Universal Mutator is simply the best available tool.   We also our results to answer a set of research questions that consider the utility of our method:

\begin{itemize}
\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?  That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between programs, or does the number of mutants killed simply reflect the ``verbosity'' of each tool?
\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?  Can we confirm results from other evaluation methods, or informal opinion?
\item {\bf RQ4:}  Are mutation scores for clean programs higher (more mutants killed) than those for programs for which tools report findings before mutation?
\item {\bf RQ5:}  Are killed mutants flagged due to real faults, where the finding is related to the introduced fault?  That is to say, is the evaluation based on meaningful semantic changes some (but not all) tools detect?
\item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
 different tools are good at/bad at, and extend this information to improve tools by addressing identified weaknesses?  How does this compare to using mutants that have not been prioritized?
  \end{itemize}

@@

\begin{figure}
  \includegraphics[width=0.6\columnwidth]{solidity.png}
  \caption{Mutants killed by Solidity static analysis tools.}
  \label{fig:solidityvenn}
\end{figure}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
    Tool & Mean & Median & Mean & Median & Ratio\\
    \hline
    \hline
    Slither & 2.37 & 1.0 & 0.09 & 0.09 & 0.038 \\
    \hline
    SmartCheck & 1.89 & 1.0 & 0.05 & 0.05 & 0.026 \\
    \hline
    Securify & 24.65 & 17.0 & 0.03 & 0.02 &  0.001 \\
    \hline
  \end{tabular}
  \caption{Solidity tool results over all contracts.}
  \label{tab:scoresolidity}
\end{table}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (3)} \\
    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
    Tool & Contracts & Mean & Median & Mean & Median\\
    \hline
    \hline
    Slither & 39 & 0.11 & 0.11 & 0.09 & 0.09 \\
    \hline
    SmartCheck & 27 & 0.03 & 0.01 & 0.03 & 0.00 \\
    \hline
    Securify & 5 & 0.00 & 0.00 & 0.00 & 0.00 \\
    \hline
  \end{tabular}
  \caption{Solidity tool clean contract results.}
  \label{tab:cleansolidity}
\end{table}

Figure \ref{fig:solidityvenn} shows the mutants killed by the Solidity analysis tools.  Tables \ref{tab:scoresolidity} and \ref{tab:cleansolidity} provide numeric details of the results, including the \emph{ratio} for each tool, adjusting its mutation scores by its' general tendency to produce findings.  First, a user examining these results would suspect that Slither and SmartCheck are both useful tools, and should likely both be applied in a high-risk security-sensitive context like smart contract development.  Second, a user might suspect that the large number of findings produced, and smaller number of mutants killed, for Securify, mean that including Securify in the static analysis tool stable is a more difficult decision.  On the one hand, Securify does detect nearly as many mutants it alone can identify as SmartCheck.  The large number of findings, and very bad mutant ratio, however, lead us to suspect that many of these ``detected'' mutants are false positives (or, at least, that the problem is not the one Securify identifies).  Extracting the signal from Securify's noise will be difficult.  We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 6 days sequential compute time for Slither and 3 days for SmartCheck, i.e., about 5-15 seconds per mutant for both tools), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.

@@

Based on the differential mutation analysis, we identified three low-hanging fruit to improve the performance of Slither.  We chose Slither in part because it seems to have a better underlying intermediate language and analysis engine, and thus is likely to produce better results for the same rule than the other tools.  The process was simple.  First, we produced a list of all mutants killed by either SmartCheck or Securify, but not killed by Slither.  We then applied the prioritization method based on the FPF algorithm and the distance metric described in Section \ref{sec:prioritizing}, and examined the mutants in rank order.  Many of the mutants were difficult to identify as true or false positives, absent context.  Some opportunities for enhancement were clear, but seemed likely to require considerable effort to implement without producing a large number of false positives.  For example, Securify often detected when an ERC20 token contract's guard preventing making the special 0x0 address the owner of a contract was removed, and issued the error {\tt Violation for MissingInputValidation}. Detecting such missing guards is probably useful, but formulating a way to do it without producing false positives is non-trivial.  We wanted to show that mutants could identify \emph{useful} but \emph{easy to implement} missing detectors.  Examining the first few mutants, we identified three such, based on mutants killed by either SmartCheck or Securify, or both:

@@

All three of these detectors were submitted as PRs, vetted over an internal benchmark set of contracts used by the Slither developers to evaluate new detectors, and accepted for release in the public version of Slither.  All three detectors produce some true positives (actual problems, though not always exploitable) in benchmark contracts, have acceptably low false positive rates, and were deemed valuable enough to include as non-informational (medium severity) detectors.  The first mutants in prioritized rank exhibiting the issues, shown above, were the 2nd, 9th, and 12th non-statement-deletion mutants ranked for SmartCheck, out of over 800 such mutants.  Using our prioritization, it was possible to identify these issues by examining fewer than 20 unkilled mutants.  Without prioritization, on average a developer would have to look at more than 200, 80, and 400 mutants, respectively, to find instances of these problems.  Interestingly, the very fact that these instances are ``needles in a haystack'' among the mutants not killed by Slither means that the results in Figure \ref{fig:solidityvenn} and Tables \ref{tab:scoresolidity} and \ref{tab:cleansolidity} are almost unaltered by our improvements to Slither: our analysis is fairly robust to modest tool improvements, unless added detectors account for a large number of mutants not detected by the tool.  Adding such detectors will also only improve mutation ratio if they do not add many false positives.  Substantial changes in results therefore require adding very effective (for mutants) detectors that seldom trigger for correct code (or at least trigger much less than for mutants).  ``Cheating'' with respect to a mutation benchmark is thus, we hope, difficult without actual major tool improvements.

@@

\begin{figure}
  \includegraphics[width=0.6\columnwidth]{java.png}
  \caption{Mutants killed by Java static analysis tools.}
  \label{fig:javavenn}
\end{figure}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
    Tool & Mean & Median & Mean & Median & Ratio\\
    \hline
    \hline
    SpotBugs & 28.93 & 14.00 & 0.07 & 0.07 & 0.002 \\
    \hline
    PMD & 53.73 & 32.00 & 0.07 & 0.07 & 0.001 \\
    \hline
    Infer & 11.60 & 3.00 & 0.00 & 0.00 &  0.000 \\
    \hline
  \end{tabular}
  \caption{Java tool results over all projects.}
  \label{tab:scorejava}
\end{table}

\begin{table}
  \begin{tabular}{l|r|r|r}
    & & \multicolumn{2}{|c|}{} \\
    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} \\
    Tool & Projects & Mean & Median \\
    \hline
    \hline
    SpotBugs & 3 & 0.05 & 0.06 \\
    \hline
    PMD & 0 & N/A & N/A \\
    \hline
    Infer & 6 & 0.00 & 0.00 \\
    \hline
  \end{tabular}
  \caption{Java tool clean project results.}
  \label{tab:cleanjava}
\end{table}

Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorejava} and \ref{tab:cleanjava} provide numeric results for projects and clean projects, respectively.

In terms of {\bf RQ1}, the raw kills results suggest there is considereable value in running both SpotBugs and PMD.  Both produce a large number of unique detections, though PMD produces about 20\% more than SpotBugs.
Infer on the other hand, is only able to detect two mutants, but these are unique.  Unfortunately, these mutants only served to indicate that we were actually running Infer in such a way that it \emph{could} detect problems:  both were concurrency warnings, but the code change in both cases was simply removing a (semantically inoperative) {\tt \@Override} annotation.  It may be that the diff sizes in our code were simply too small for Infer's approach.

{\bf RQ2} is also answered in the affirmative.  While the raw kills for SpotBugs are not as good as for PMD, it produced fewer findings, giving it a mutant ratio approximately twice that of PMD.

@@

For Java, again, the large majority of mutants we inspected for {\bf RQ5}, where at least one tool detected the mutant and at least one tool did not detect it were definitely meaningful.  In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null} that would clearly introduce potential null pointer exceptions (NPEs), and such a possible NPE was the produced finding.  While the differences between Solidity tools were often due to different detectors, the Java differences seemed mostly rooted in analysis engine methods; all tools aim to warn about potential NPEs.  Because the number of mutants we could examine and understand was smaller, we are less confident in making a probabilistic estimate than with Solidity, but it was clear the basis of the comparison was primarily realistic faults, which some tools detected and others did not.

@@

\begin{figure}
  \includegraphics[width=0.6\columnwidth]{python.png}
  \caption{Mutants killed by Python static analysis tools.}
  \label{fig:pythonvenn}
\end{figure}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
    Tool & Mean & Median & Mean & Median & Ratio\\
    \hline
    \hline
    Pylint & 2.24 & 1.00 & 0.22 & 0.20 & 0.097 \\
    \hline
    pyflakes & 2.4 & 1.00 & 0.03 & 0.00 & 0.012 \\
    \hline
    PyChecker & 0.56& 0.00 & 0.10 & 0.00 &  0.183 \\
    \hline
  \end{tabular}
  \caption{Python tool results over all projects.}
  \label{tab:scorepython}
\end{table}

\begin{table}
  \begin{tabular}{l|r|r|r}
    & & \multicolumn{2}{|c|}{} \\
    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} \\
    Tool & Projects & Mean & Median \\
    \hline
    \hline
    Pylint & 0 & N/A & N/A \\
    \hline
    pyflakes & 12 & 0.02 & 0.00 \\
    \hline
    PyChecker & 21& 0.01 & 0.00 \\
    \hline
  \end{tabular}
  \caption{Python tools clean project results.}
  \label{tab:cleanpython}
\end{table}

Figure \ref{fig:pythonvenn} shows the mutants killed by the Python analysis tools, and Tables \ref{tab:scorepython} and \ref{tab:cleanpython} provide numeric results for projects and clean projects, respectively.

@@

The mutant ratios {\bf RQ2} suggest that PyChecker might be better than pyflakes, in fact, and better than Pylint, because while it killed fewer mutants, it also produced fewer findings by far, overall.  PyChecker would be a good choice for quick and dirty analysis of code on the fly, perhaps, in that it had excellent mutation scores for some programs (hence a high mean) but is relatively ``quiet'' compared to the other tools.  However, its very low median mutation score is troubling; it may simply not work for some code.  A combination of pyflakes and Pylint seems safer, though adding the relatively low-findings PyChecker seems likely to have little cost.  For Python, we see that all of our measures are required to get a good picture of the tools.

>>>>

Our primary experimental results are a set of comparisons of tools using our method, for three languages: Solidity (the most popular language for smart contracts), Java, and Python.  We used the Universal Mutator tool for all experiments; for Solidity and Python, we believe the Universal Mutator is simply the best available tool.  For Java, PIT \cite{pitmut} is more popular, but does not produce source-level mutants, needed for PMD and for manual inspection of results.
We used our results to answer a set of research questions:

\begin{itemize}[labelsep=3pt,leftmargin=12pt]
\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?  That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar?
\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between static analysis tools?
\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?  Can we independently confirm our results?
\item {\bf RQ4:}  Do tools detect more mutants in programs for which they produce no warnings, initially?
\item {\bf RQ5:}  Are mutants distinguishing tools usually flagged due to real faults, where the finding is related to the introduced fault; that is, are our results usually \emph{meaningful}?  
\item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
 different tools are good at/bad at, and use this information to improve tools?  How does this compare to using mutants that have \emph{not} been prioritized?
  \end{itemize}

@@

\begin{figure}
  \includegraphics[width=0.4\columnwidth]{solidity.png}
  \caption{Mutants killed by Solidity static analysis tools.}
  \label{fig:solidityvenn}
\end{figure}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
    Tool & Mean & Median & Mean & Median & Ratio\\
    \hline
    \hline
    Slither & 2.37 & 1.0 & 0.09 & 0.09 & 0.038 \\
    Clean (39) & - & - & 0.11 & 0.11 & -\\
    \hline
    SmartCheck & 1.89 & 1.0 & 0.05 & 0.05 & 0.026 \\
    Clean (27) & - & - & 0.03 & 0.01 & - \\
    \hline
    Securify & 24.65 & 17.0 & 0.03 & 0.02 &  0.001 \\
    Clean (5) & - & - & 0.00 & 0.00 & - \\
    \hline

  \end{tabular}
  \caption{Solidity tool results over all contracts.}
  \label{tab:scoresolidity}
\end{table}

Figure \ref{fig:solidityvenn} shows the mutants killed by the Solidity analysis tools.  Table \ref{tab:scoresolidity} provides numeric details of the results, including the \emph{ratio} for each tool, adjusting its mutation scores by its' general tendency to produce findings.  The second row for each tool shows the number of contracts for which it reported no findings, and the mutation scores over those contracts, only. A user examining these results would suspect that Slither and SmartCheck are both useful tools, and should likely both be applied in a high-risk security-sensitive context like smart contract development.  A user might also suspect that the large number of findings produced, and smaller number of mutants killed, for Securify, means that whether to apply Securify is a more difficult decision.  On the one hand, Securify does detect nearly as many mutants it alone can identify as SmartCheck.  The large number of findings, and very bad mutant ratio, however, lead us to suspect that many of these ``detected'' mutants are false positives (or, at least, that the problem is not the one Securify identifies).  Extracting the signal from Securify's noise will be difficult.  We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 6 days sequential compute time for Slither and 3 days for SmartCheck, i.e., about 5-15 seconds per mutant for both tools), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.

@@

Based on the differential mutation analysis, we identified three low-hanging fruit to improve the performance of Slither.  We chose Slither in part because it seems to have a better underlying intermediate language and analysis engine, and thus is likely to produce better results for the same rule than the other tools.  The process was simple.  First, we produced a list of all mutants killed by either SmartCheck or Securify, but not killed by Slither.  We then applied the prioritization method based on the FPF algorithm and the distance metric described in Section \ref{sec:prioritizing}, and examined the mutants in rank order.  Many of the mutants were difficult to identify as true or false positives, absent context.  Some opportunities for enhancement were clear, but seemed likely to require considerable effort to implement without producing a large number of false positives.  For example, Securify often detected when an ERC20 token contract's guard preventing making the special 0x0 address the owner of a contract was removed, and issued the error {\tt Violation for MissingInputValidation}. Detecting such missing guards is probably useful, but doing so without producing false positives is non-trivial.  We wanted to show that mutants could identify \emph{useful} but \emph{easy to implement} missing detectors.  Examining only a few mutants, we identified three:

@@

All three of these detectors were submitted as PRs, vetted over an internal benchmark set of contracts used by the Slither developers to evaluate new detectors, and accepted for release in the public version of Slither.  All three detectors produce some true positives (actual problems, though not always exploitable) in benchmark contracts, have acceptably low false positive rates, and were deemed valuable enough to include as non-informational (medium severity) detectors.  The first mutants in prioritized rank exhibiting the issues, shown above, were the 2nd, 9th, and 12th non-statement-deletion mutants ranked for SmartCheck, out of over 800 such mutants.  Using our prioritization, it was possible to identify these issues by examining fewer than 20 unkilled mutants.  Without prioritization, on average a developer would have to look at more than 200, 80, and 400 mutants, respectively, to find instances of these problems.  Interestingly, the very fact that these instances are ``needles in a haystack'' among the mutants not killed by Slither means that the results in Figure \ref{fig:solidityvenn} and Table \ref{tab:scoresolidity} are almost unaltered by our improvements to Slither: our analysis is fairly robust to modest tool improvements, unless added detectors account for a large number of mutants not detected by the tool.  Adding such detectors will also only improve mutation ratio if they do not add many false positives.  Substantial changes in results therefore require adding very effective (for mutants) detectors that seldom trigger for correct code (or at least trigger much less than for mutants).  ``Cheating'' with respect to a mutation benchmark is thus, we hope, difficult without actual major tool improvements.

@@

\begin{figure}
  \includegraphics[width=0.4\columnwidth]{java.png}
  \caption{Mutants killed by Java static analysis tools.}
  \label{fig:javavenn}
\end{figure}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
    Tool & Mean & Median & Mean & Median & Ratio\\
    \hline
    \hline
    SpotBugs & 28.93 & 14.00 & 0.07 & 0.07 & 0.002 \\
    Clean (3) & - & - & 0.05 & 0.06 - \\
    \hline
    PMD & 53.73 & 32.00 & 0.07 & 0.07 & 0.001 \\
    Clean (0) & - & - & - & - & - \\
    \hline
    Infer & 11.60 & 3.00 & 0.00 & 0.00 &  0.000 \\
    Clean (6) & - & - & 0.00 & 0.00 \\
    \hline
  \end{tabular}
  \caption{Java tool results over all projects.}
  \label{tab:scorejava}
\end{table}

Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Table \ref{tab:scorejava} provides numeric results for projects and clean projects, respectively.

In terms of {\bf RQ1}, the raw kills results suggest there is considereable value in running both SpotBugs and PMD.  Both produce a large number of unique detections, though PMD produces about 20\% more than SpotBugs.
Infer on the other hand, is only able to detect two mutants, but these are unique.  Unfortunately, these mutants only served to indicate that we were actually running Infer in such a way that it \emph{could} detect problems:  both were concurrency warnings, but the code change in both cases was simply removing a (semantically inoperative) {\tt \@Override} annotation.  It may be that the diff sizes in our code were simply too small for Infer's approach.
{\bf RQ2} is also answered in the affirmative.  While the raw kills for SpotBugs are not as good as for PMD, it produced fewer findings, giving it a mutant ratio approximately twice that of PMD.

@@

For Java, again, the large majority of mutants we inspected for {\bf RQ5}, where at least one tool detected the mutant and at least one tool did not detect it were definitely meaningful.  In particular, for Java, the large majority of mutants involved either deleted method calls or changes to conditionals (e.g., {\tt == null} to {\tt != null}) that would clearly introduce potential null pointer exceptions (NPEs), and such a possible NPE was the produced finding.  While the differences between Solidity tools were often due to different detectors, the Java differences seemed mostly rooted in analysis engine methods; all tools aim to warn about potential NPEs.  Because the number of mutants we could examine and understand was smaller, we are less confident in making a probabilistic estimate than with Solidity, but it was clear the basis of the comparison was primarily realistic faults, which some tools detected and others did not.

@@

\begin{figure}
  \includegraphics[width=0.4\columnwidth]{python.png}
  \caption{Mutants killed by Python static analysis tools.}
  \label{fig:pythonvenn}
\end{figure}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
    Tool & Mean & Median & Mean & Median & Ratio\\
    \hline
    \hline
    Pylint & 2.24 & 1.00 & 0.22 & 0.20 & 0.097 \\
    Clean (0) & - & - & - & - & - \\
    \hline
    pyflakes & 2.4 & 1.00 & 0.03 & 0.00 & 0.012 \\
    Clean (12) & - & - & 0.02 & 0.00 & - \\
    \hline
    PyChecker & 0.56& 0.00 & 0.10 & 0.00 &  0.183 \\
    Clean (21) & - & - & 0.01 & 0.00 \\
    \hline
  \end{tabular}
  \caption{Python tool results over all projects.}
  \label{tab:scorepython}
\end{table}

Figure \ref{fig:pythonvenn} shows the mutants killed by the Python analysis tools, and Table \ref{tab:scorepython} provides numeric results for projects and clean projects, respectively.

@@

The mutant ratios {\bf RQ2} suggest that PyChecker might be better than pyflakes, in fact, and better than Pylint, because while it killed fewer mutants, it also produced fewer findings by far, overall.  PyChecker would be a good choice for quick and dirty analysis of code on the fly, perhaps, in that it had excellent mutation scores for some programs (hence a high mean) but is relatively ``quiet'' compared to the other tools.  However, its very low median mutation score is troubling; it may simply not work for some code.  A combination of pyflakes and Pylint seems safer, though adding the relatively low-findings PyChecker seems likely to have little cost.  For Python, all of our measures are required to get a good picture.

### 2020-04-20 introduction.tex
ok

Mutation testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical} (or mutation analysis, the term we will use in this work, for reasons that will become clear) uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is mostly correct, most small changes will therefore introduce a fault.  For the most part, mutation analysis has been used to evaluate test suites by computing a mutation score, the fraction of mutants the suite detects, or ``kills''.  Most such use has been in research efforts, rather than practical testing efforts, though there has been sporadic use by interested developers.
Groce et al. \cite{groce2015verified,groce2018verified} proposed examining individual mutants that survive a rigorous testing and verification effort to detect and correct weaknesses in such efforts.  The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and a widely used Python file system.  Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, due to scale.

Combining a differential approach (not differential \emph{testing}, precisely, in that individual differences are not always worth inspecting) and mutation analysis, however, offers a novel way to compare static analysis tools, one useful to users wishing to select a good tool or set of tools, researchers interested in, e.g., the impact of precision/soundness trade-offs or different intermediate languages, and developers of static analysis tools hoping to improve their tools by systematic examination of code changes that are likely fault-inducing but not detected.

@@

\begin{figure}
  \includegraphics[width=0.5\columnwidth]{example.png}
  \caption{Mutants killed by three static analysis tools.}
  \label{fig:examplevenn}
\end{figure}

@@

Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool, and thus known to be in-principle detectable, but not detected by the tool to be improved.  Improving tools by adding detectors is useful because, even if all tools had the same set of detectors, they would not all report the same bugs; difference choices in intermediate language and tradeoffs made to avoid false positives would make the use of multiple tools with shared detectors essential for thorough analysis.  And, of course, if one tool does have a much better underlying engine, it is beneficial to users that the ``best'' tool incorporate all detection rules.

Of course, any faults in code, not just mutants, could serve this purpose.  But, again, the automatic nature of mutation generation, and the presumption that the mutation is indeed a fault, is useful.  Moreover, because mutants follow syntactic patterns, searching for similar mutants/faults the tool to be improved does not detect is much easier than with arbitrary faults, and can be partly automated.   However, as with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.  We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases to find distinct faults \cite{PLDI13,distMut}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting mutants not identified for similar reasons.

@@

\begin{itemize}
\item We propose a differential approach to comparing static analysis tools based on the insight that program mutants provide an automated source of simple, easy-to-understand program changes that are likely faults.
\item We propose a definition of mutant killing that works well in a static analysis context.
\item We introduce a simple scheme for prioritizing mutants that helps users understand the results of analysis and guide efforts to improve, rather than simply compare, tools.
\item We apply our method to an extensive, in-depth comparison of three Solidity smart contract analysis tools, and show how prioritization allowed us to easily identify (and build) three new detectors for the most effective of these tools.
\item We further provide results for comparisons of popular Java and Python static analysis tools, showing the general usefulness of our methods, and giving a new picture of the comparative effectiveness of these tools.
\end{itemize}

>>>>

Mutation analysis\cite{jia2011analysis,demillo1978hints,budd1980theoretical} uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is mostly correct, such changes will often introduce a fault.  For the most part, mutation analysis has been used to evaluate test suites by computing a mutation score, the fraction of mutants the suite detects, or ``kills''.  Most such use has been in research efforts, rather than practical testing efforts, though there has been sporadic use by interested developers.
Groce et al. \cite{groce2015verified,groce2018verified} proposed examining individual mutants that survive a rigorous testing and verification effort to detect and correct weaknesses in such efforts.  The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and a widely used Python file system.  Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, due to scale.

Combining a differential approach and mutation analysis offers a novel way to compare static analysis tools, one useful to users wishing to select a good tool or set of tools, to researchers interested in the impact of precision/soundness trade-offs or different intermediate languages, and to developers of static analysis tools hoping to improve their tools by examining missed faults.

@@

\begin{figure}
  \includegraphics[width=0.4\columnwidth]{example.png}
  \caption{Mutants killed by three static analysis tools.}
  \label{fig:examplevenn}
\end{figure}

@@

Comparing mutant results also leads to the idea of \emph{improving} static analysis tools by examining mutants detected by another tool, and thus known to be in-principle detectable, but not detected by the tool to be improved.  Improving tools by adding detectors is useful because, even if all tools had the same set of detectors, they would not all report the same bugs; different choices in intermediate language and tradeoffs made to avoid false positives may make the use of multiple tools with similar detectors essential for thorough analysis.  And if one tool simply has a superior engine, it is beneficial to users that the ``best'' tool incorporate \emph{all} detection rules.

Of course, any faults in code, not just mutants, could serve this purpose.  But, again, the automatic nature of mutation generation, and the presumption that the mutation is indeed a fault, is useful.  Moreover, because mutants follow syntactic patterns, searching for similar mutants/faults the tool to be improved does not detect is much easier than with arbitrary faults, and can be partly automated.   However, as with efforts to improve test suites, manually searching through all mutants can be an onerous task, especially for large-scale evaluations.  We therefore also introduce the idea of prioritizing mutants, using a Furthest-Point-First \cite{Gonzalez} algorithm and distance metric inspired by previous work on helping developers sort through failing test cases to find distinct faults \cite{PLDI13,distMut}, to help static analysis tool developers find interesting patterns without wading through numerous uninteresting mutants.

@@

\begin{itemize}[labelsep=3pt,leftmargin=12pt]
\item We propose a differential approach to comparing static analysis tools based on the insight that program mutants provide an automated source of simple, easy-to-understand program changes that are likely faults.
\item We propose a definition of mutant killing for static analysis.
\item We introduce a simple scheme for prioritizing mutants that helps users understand and use the results of analysis.
\item We apply our method to an extensive, in-depth comparison of three Solidity smart contract analysis tools, and show how prioritization allowed us to easily identify (and build) three new detectors for the most effective of these tools.
\item We further provide results for comparisons of popular Java and Python static analysis tools, showing the general usefulness of our methods, and giving a new picture of the comparative effectiveness of these tools.
\end{itemize}
